### 2015-09-07 paper/highlights.tex

\documentclass{article}

\begin{document}

\noindent Highlights:, \\

\vspace{0.25in}

\noindent A spatial agent-based model examines T cell search in an influenza infected lung. \\

\noindent Experimental data show chemokines CXCL10 and CCL5 are stimulated by infection. \\

\noindent Rapidly expanding plaques of infection create challenges for chemotaxing T cells. \\

\noindent Sensitivity analysis isolates the effects of each parameter in the model. \\

\noindent Model parameters are used to determine an upper bound on T cell search time. \\

\end{document}